Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex

We assume that the domain of the observer model is further discretised, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, where $*$ represents the appropriate
dimension for each matrix.
New problems arise due to the FWL effects on these matrices.  
The canonical form of the resulting model is expressed by:
%
{\setlength{\belowdisplayskip}{2pt}
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right]%\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
}
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_c=&\left[\begin{array}{cc}
\begin{array}{ccc}k_{n-1}-a_{n-1}&\cdots&k_1-a_1\end{array}&k_0-a_0\\
\mat{I}&\begin{array}{c}0\\ \vdots \\ 0\end{array}
\end{array}\right]&
\mat{A}_o=&\left[\begin{array}{cc}
\begin{array}{c}l_{n-1}-a_{n-1}\\ \vdots\\ l_1-a_1\end{array}&\mat{I}\\
l_0-a_0&\begin{array}{ccc}0&\cdots&0\end{array}
\end{array}\right]
\end{align*}
}

where $\vec{x}'$ and $\vec{x}_e'$ are the state of the system and the error of the observer evaluated in the transformed domain, 
$\mat{A}_{c}$ the closed loop dynamics of the controller in canonical form, $\mat{A}_{o}$ the closed loop dynamics of the observer in canonical form, $\mat{B}_{o}$ the feedback gain, 
 $\mat{B}'$ the noise sensitivity, and $\mat{A}_{e}$ and $\mat{B}_{e}$ the error caused by the FWL representation.
%and we have introduced the following terms: 
%{\setlength{\abovedisplayskip}{2pt}
%\setlength{\belowdisplayskip}{2pt}
%\begin{align*}
%\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
%\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
%\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
%\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
%\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
%\mat{C}_{e}=&\mat{C}-\mat{C}_{\langle C \rangle}
%\end{align*}
%}
%{\setlength{\abovedisplayskip}{0pt}
%\begin{align*}
%\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
%\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
%\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
%\end{align*}
%}
Here $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations (details are in Appendix~\ref{sec:quantization-noise}).

In a continuous time domain, the characteristic polynomial
of the closed loop dynamics is defined by the characteristic polynomials of~$\mat{A}_c$
and~$\mat{A}_o$.  The introduction of the FWL errors causes this to no longer
be true.  We must therefore find a suitable solution to maintain this separation.

Gershgorin disks~\cite{van1996matrix} allow us to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda}_{oi}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

In the second stage, we take the complex eigenvalue
with the largest angle ($\max\{|\tan^{-1}\left(\frac{im(\lambda)}{re(\lambda)}\right)| \mid \lambda \in \Lambda\}$)
and reduce its imaginary component only  ($re(\lambda) + f\, im(\lambda) \models \phi$).
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and has a
smaller rotation, which facilitates meeting the safety specification $\phi_\mathit{safety}$.
%(we note that real convergent eigenvalues are trivially safe).

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 1.1\,s &   3.8\,s & 2\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 7.1\,s  &   18.0\,s & 9\\
3  & Inverted Pendulum & 4 & 100 &   & \xmark &\xmark & 20.0\,s  &  348.1\,s & 19\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 3.0\,s  &  9.5\,s & 7 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 33 &0.4\,s  &   4.9\,s & 6\\
6  & Suspension        & 4 & 100 & & & \xmark& 14.8\,s&166.9\,s  & 26 \\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11&3.1\,s  &   16.8\,s & 8\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, and Median times for standard CEGIS vs CEGIS-OR on multiple benchmarks. t.o.~indicates timeouts. \note{It is unclear what the difference is btw t.o. used in the third column header, and that used within the table - the latter is corresponding to a cross?} \note{recall what a model order is.}}
\end{table}

Using the described technique we managed to synthesize controllers for \note{X} benchmarks, with a median synthesis time of $18.8$ seconds.
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control engineers.
There are cases for which the system fails to find a controller.  We use a
relatively short timeout, so it is possible that longer timeouts would instead lead to a
valid solution. However, as stated before, the solution is sound but incomplete,
so it is impossible to tell if a solution exists in these cases.

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority of
our benchmarks, we observe that our approach is significantly faster than
standard CEGIS. We further note that the number of times standard CEGIS times
out is much larger than CEGIS-OR which results in more found solutions.
The magnitude of the improvement is however, largely dependent on the shape of the problem.

More interesting cases appear when we increase the order, showing initially that
benchmarks that timed-out using standard CEGIS are now running in under 2
minutes using CEGIS-OR. This is the result of removing several multiplications which are
known to be difficult for SAT solvers.  Furthermore, when we introduce Optimization
Refinement there is a further improvement up to an order of magnitude.  This second
improvement is due to the elimination of counterexamples that would have been
created without the refinement. Since further synthesis require increasingly
complex models which are linear in the number of multiplications, eliminating 
these results in large improvements.  In cases when the initial candidate is
suitable, we gain nothing with the refinement and therefore find similar times.
Improvements using this approach will therefore be observed in higher order
systems or when the safety specification is tight and the original loop would
require multiple counterexamples.

>>>>

We assume that the domain of the observer model is further discretised, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, where $*$ represents the appropriate
dimension for each matrix.
New problems arise due to the FWL effects on these matrices.  
The canonical form of the resulting model is expressed by:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right]%\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where $\vec{x}'$ and $\vec{x}_e'$ are the state of the system and the error of the observer evaluated in the transformed domain, 
$\mat{A}_{c}$ the closed loop dynamics of the controller, $\mat{A}_{o}$ the closed loop dynamics of the observer, $\mat{B}_{o}$ the feedback gain, 
 $\mat{B}'$ the noise sensitivity, and $\mat{A}_{e}$ and $\mat{B}_{e}$ the error caused by the FWL representation.
%and we have introduced the following terms: 
%{\setlength{\abovedisplayskip}{2pt}
%\setlength{\belowdisplayskip}{2pt}
%\begin{align*}
%\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
%\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
%\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
%\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
%\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
%\mat{C}_{e}=&\mat{C}-\mat{C}_{\langle C \rangle}
%\end{align*}
%}
%{\setlength{\abovedisplayskip}{0pt}
%\begin{align*}
%\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
%\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
%\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
%\end{align*}
%}
Here $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations (details are in Appendix~\ref{sec:quantization-noise}).

\note{the remainder of this section is not self contained and totally unclear.}

In a continuous time domain, the characteristic polynomial
of the closed loop dynamics is defined by the characteristic polynomials of~$\mat{A}_c$
and~$\mat{A}_o$.  The introduction of the FWL errors causes this to no longer
be true.  We must therefore find a suitable solution to maintain this separation.
Gershgorin disks~\cite{van1996matrix} allow us to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda}_{oi}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda}. 
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda}. 
\end{equation}

@@

In the second stage, we take the complex eigenvalue
with the largest angle ($\max\{|\tan^{-1}\left(\frac{im(\lambda)}{re(\lambda)}\right)| \mid \lambda \in \Lambda\}$)
and reduce its imaginary component only  ($re(\lambda) + f\, im(\lambda) \models \phi$).
\note{notation again.}
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and has a
smaller rotation, which facilitates meeting the safety specification $\phi_\mathit{safety}$.
%(we note that real convergent eigenvalues are trivially safe).

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 1.1\,s &   3.8\,s & 2\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 7.1\,s  &   18.0\,s & 9\\
3  & Inverted Pendulum & 4 & 100 &   & \xmark &\xmark & 20.0\,s  &  348.1\,s & 19\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 3.0\,s  &  9.5\,s & 7 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 33 &0.4\,s  &   4.9\,s & 6\\
6  & Suspension        & 4 & 100 & & & \xmark& &  & \xmark\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11&3.1\,s  &   16.8\,s & 8\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, and Median times for standard CEGIS vs CEGIS-OR on multiple benchmarks.  the t.o. column ~indicates number of timeouts. \note{It is unclear what the difference is btw t.o. used in the third column header, and that used within the table - the latter is corresponding to a cross?} \note{recall what a model order is.}}
\end{table}

Using the described technique we managed to synthesize controllers for \note{X} benchmarks, with a median synthesis time of $18.8$ seconds.
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control engineers.
There are cases for which the system fails to find a controller.  We use a
relatively short timeout, so it is possible that longer timeouts would instead lead to a
valid solution. However, as stated before, the solution is sound but incomplete,
so it is impossible to tell if a solution exists in these cases. 
\note{I don't think this is stated before so explain it here.}

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority of
our benchmarks, we observe that our approach is significantly faster than
standard CEGIS. We further note that the number of times standard CEGIS times
out is much larger than CEGIS-OR which results in more found solutions.e

%% We observe that the magnitude of the improvement is however, largely dependent on the shape of the problem.

Overall, we observed more improvements with CEGIS-OR for
higher order systems or systems where the safety
specification is tight, meaning that the standard CEGIS loop would require many iterations.
%  We obtain the most improvements for the models with large orders?}
%More interesting cases appear when we increase the order, showing
For instance, benchmark X, which timed-out when using standard CEGIS,
is running under 2 minutes using CEGIS-OR. This is the result of
removing several multiplications which are known to be difficult for
SAT solvers \note{I'm not sure readers will relate to where this is happening in the paper. is it the fact that you use that canonical form? then say that. however, i would have thought you do that in
both approaches..}.
Another factor that brings an improvement of one order of
magnitude is the Abstraction Refinement and Optimization in CEGIS-OR
\note{on average/an improvement of up to an order of magnitude/for some specific case?}.
While abstraction refinement enables us to start with coarse abstractions
that are only refined on demand, 
the optimization
allows to locally improve candidate solutions and avoid entire iterations of the CEGIS loop
(which would have also required more refined abstractions).
%% elimination of counterexamples that
%% would have been created without the refinement. Since further
%% synthesis require increasingly complex models which are linear in the
%% number of multiplications, eliminating these results in large
%% improvements.
While some systems might still require a very pecise
abstraction resulting in many refinements
(this is generally the cause of time-outs in CEGIS-OR),
others allow us to find solutions with rather coarse
abstractions (this is reflected in the smaller minimum
times obtained with CEGIS-OR).

%% Additionally, as abstraction refinement allows us to start with coarse initial
%% abstraction, CEGIS-OR has smaller minimum times than CEGIS, obtained
%% when the initial coarse abstractions were sufficient to synthesise a
%% solution.

%% In cases when the initial candidate is suitable, we
%% gain nothing with the refinement.
%% Overall, we observed more improvements with CEGIS-OR for
%% higher order systems or systems where the safety
%% specification is tight, where the standard CEGIS loop would require many iterations.
%multiple counterexamples.
